---
title: "Assignment 2"
author: "Orli Khaimova"
date: "9/5/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#install.packages("RMySQL")
```

```{r}
library(RMySQL)
library(dplyr)
```
### Libraries Used
I used the *RMySQL* and *dplyr* libraries to incorporate MySQL and to perform other functions.

```{r mysql setup}
db <- dbConnect(MySQL(), user='root', password=.rs.askForPassword("Enter password:"), 
                dbname='movies', host='localhost')

summary(db)

#lists the tables in the database
dbListTables(db)

#assigning variables to each table
responders <- dbReadTable(db, "responders")
ratings <- dbReadTable(db, "ratings")
```
### MySQL Setup
First, I set up the connection to the database. Instead of writing the password, I programmed it so that it would instead prompt the user to enter the password. This way the password is not shared when others are viewing. I assigned the two different tables to different variables so that it would be use to refer to later. Strangely, when I set it up to load from file, "\r" was added to each name.

```{r table making}
#merging the two tables
MovieRatings <- left_join(responders, ratings, by = c("responder_id" = "survey_id"))

## removing the id variable
MovieRatings <- subset(MovieRatings, select = -c(responder_id))

## renaming variable
MovieRatings <- rename(MovieRatings, "name" = "responder")

MovieRatings
```
### Table Making
I merged the two tables by using a left join. I removed the `id` variable so that it would be easier to view. I also removed the id variable and renamed another variable. 

```{r analysis}
summary(MovieRatings)

MovieRatings %>% summarize_if(is.numeric, mean, na.rm = TRUE) 
```

### Short Analysis
Based on these results, Joker had the highest ratings with Cats having the least. 