---
title: "Week 7: Working with XML and JSON in R"
author: "Orli Khaimova"
date: "10/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(rvest)
library(xml2)
library(XML)
library(RCurl)
library(tidyverse)
library(httr)
library(knitr)
library(RJSONIO)
```

### Task Description

Separately create three files which store the book’s information in HTML , XML, and JSON formats containing book information and include the title, authors, and two or three
other attributes that you find interesting. With R and using any packages, load information from the three sources into separate data frames. Check if all three are identical.

### Overview of Approach

I created each file by hand and uploaded it to *Github* so that it can be read and reproduced. I then converted each one into a data frame. Lastly, I compared to see if they
were identical. 


### Results{.tabset}

#### HTML

```{r HTML}
books_html <- read_html("https://raw.githubusercontent.com/okhaimova/DATA-607/master/Week7/books.html") %>%
  html_nodes("table") %>% 
  html_table() %>% 
  as.data.frame()

kable(books_html)
```
#### XML

```{r XML}
books_xml <- xmlParse(getURL("https://raw.githubusercontent.com/okhaimova/DATA-607/master/Week7/books.xml")) %>%
  xmlToDataFrame()
  
kable(books_xml)
```
#### JSON

Note: I had to remove row names and also replace a string of characters as one author has a German letter in their name. However, if there is unicode data in the file, you 
can just include `encoding = "UTF-8"` when loading data. I also trasposed the data frame.

```{r JSON}
books_json <- fromJSON("https://raw.githubusercontent.com/okhaimova/DATA-607/master/Week7/books.json", encoding = "UTF-8") %>%
  as.data.frame() %>%
  t() %>%
  as.data.frame() %>%
  remove_rownames()

kable(books_json)
```

### Comparison & Findings

Using the transitive property, I compared `books_html` to `books_html` and to `books_xml`. The end result was that all three data frames were identical.

```{r}
books_html == books_json && books_html == books_xml
```

