---
title: "Project 1: Data Analysis: Chess Tournament"
author: "Orli Khaimova"
date: "9/19/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

### Reading the Data
```{r}
URL <- "https://raw.githubusercontent.com/okhaimova/DATA-607/master/Project1/tournamentinfo"
tournamenttemp <- read.csv(URL, header = FALSE, sep = "|")
```

### Cleaning up the Data

After reading the data, we have to clean it up. I removed the dashes and then separated the data into two separate data frames. One consists of the pair number, player name, total points, and rounds. The second consists of the state, USCF ID, pre-rating, post-rating, and more round information. They are labeled `odd` and `even` respectively. I also split the variables further in order to analyze it and made some variables of the numeric class.
```{r}
#removing the lines with dashes
tournamenttemp <- tournamenttemp[!grepl("---", tournamenttemp[ ,1]), ]

#removing the first 2 columns
tournamenttemp <- tournamenttemp[-c(1:2), ]

#removing V11 column
tournamenttemp <- tournamenttemp[ ,-c(11)]

#splitting data frame into two
odd <- tournamenttemp[seq(1, nrow(tournamenttemp), 2), ]
even <- tournamenttemp[seq(2, nrow(tournamenttemp), 2), ]

#renaming variables in odd
colnames(odd) <- c("pair", "name", "points", 1:7)

#clean up the even data frame
#removing + renaming variables
even <- even[ ,-c(3:10)]
even <- rename(even, "state" = "V1")

#splitting columns
even <- separate(even, 2, c("uscfID", "rating" ), "/")
even <- separate(even, "rating", c("pre-rating", "post-rating"), "->")

#cleaning the ratings to make it numeric 
even$`pre-rating` <- str_extract(even$`pre-rating`, "\\d+\\d")
even$`post-rating` <- str_extract(even$`post-rating`, "\\d+\\d")

#combining data frames
tournament <- cbind(odd, even)

#cleaning up the data to make it numeric
tournament$pair <- unlist(str_extract_all(tournament$pair, "\\d+"))
tournament$points <- unlist(str_extract_all(tournament$points, "\\d.\\d"))

#changing class
tournament$`pre-rating` <- as.numeric(tournament$`pre-rating`)
tournament$points <- round(as.double(tournament$points), 2)
```

To extract the pair numbers from the rounds variables in order to match it to the opponents' pre-rating, I converted it to a matrix. I then applied a for loop which took each value in the matrix and replaced it with the opponents' pre-ratings. Afterwards, I made it numeric.
```{r}
#assigning the round columns its own variable
rounds <- tournament[4:10]

#extracting the digits from the round columns, putting and putting it in a matrix
test <- matrix(str_extract(unlist(rounds), "\\d+"), ncol = 7)

#for loop
#it goes through the `test` matrix and replaces the values with the opponents' pre-rating
for (row in 1:nrow(test))
{
  for (col in 1:ncol(test))
  {
    i <- test[row, col]
    
    if(!is.na(i))
    {
      test[row, col] <- tournament$`pre-rating`[tournament$pair == i]
    }
  }
}

#changing the character matrix into numeric
class(test) <- "numeric"

#assigning the `test` matrix back into the `rounds` data frame
rounds <- test
```

### Calculations

I then calculated the average of the opponent's pre-ratings for each player and removed the `NA` values for it to only be effective for the amount of rounds they played.
```{r}
#calculating the average of the opponents' pre-ratings
tournament$avg <- round(rowMeans(rounds, na.rm = TRUE))
```

### Output

I then created a `toutput` with only the `name`, `state`, `points`, `pre-rating`, and `avg` variables. Afterwards, it generates a CSV file with only the `toutput` data.
```{r}
#reordering columns and creating output data frame
toutput <- tournament[c("name", "state", "points", "pre-rating", "avg")]

head(toutput)

#creating csv file
write.csv(toutput, "tournament", row.names = FALSE)
```

### Calculating the players' expected scores

Using the ELO calculation, I determined each player’s expected result (number of points),
based on his or her pre-tournament rating, and the average pre-tournament rating for all of the
player’s opponents. I was able to find their probability of winning and then multiplied it by the amount of rounds they played. By doing so, I was able to calculate their expected number of points.
```{r}
toutput$win_prob <- round((1 / (1 + 10 ^ ((toutput$avg - toutput$`pre-rating`) / 400))), 4)

toutput$predicted_score <- round(toutput$win_prob * rowSums(!is.na(rounds)), 4)

toutput$diff <- round(abs(toutput$points - toutput$predicted_score), 4)

head(toutput[order(toutput$diff),], 5)  
```
### Analysis of Expected Scores

According to the calculations, Michael J Martin scored the most points relative to his expected result, followed by Hansen Song, and Bradley Shaw. 
