---
title: "Assignment 3"
author: "Orli Khaimova"
date: "9/12/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

### Problem 1

Using the 173 majors listed in fivethirtyeight.com’s [College Majors dataset] (https://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/), provide code that identifies the majors that contain either "DATA" or "STATISTICS"

```{r}
URL <- "https://raw.githubusercontent.com/fivethirtyeight/data/master/college-majors/majors-list.csv"
majors <- read.csv(URL)

majors %>% filter(str_detect(Major, "(DATA)|(STATISTICS)"))
```


### Problem 2

Write code that transforms the data below:

[1] "bell pepper"  "bilberry"     "blackberry"   "blood orange"

[5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"  

[9] "elderberry"   "lime"         "lychee"       "mulberry"    

[13] "olive"        "salal berry"

Into a format like this:

c("bell pepper", "bilberry", "blackberry", "blood orange", "blueberry", "cantaloupe", "chili pepper", "cloudberry", "elderberry", "lime", "lychee", "mulberry", "olive", "salal berry")

```{r}
fruits <- '[1] "bell pepper"  "bilberry"     "blackberry"   "blood orange"

[5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"  

[9] "elderberry"   "lime"         "lychee"       "mulberry"    

[13] "olive"        "salal berry"'

fruits <- str_extract_all( fruits, '\"[a-z]*\\s*[a-z]*\\"')

fruits <- unlist(fruits)

fruits

writeLines(fruits)


```



The two exercises below are taken from R for Data Science, 14.3.5.1 in the on-line version:


### Problem 3

Describe, in words, what these expressions will match:

* (.)\1\1
  + Any character that is the same would be repeated 3 times
* "(.)(.)\\2\\1"
  + A string with a symmetric pattern `abba`. There are two characters followed by the second character and then the first character
* (..)\1
  + A pair of characters that is repeated once such as `abab`
* "(.).\\1.\\1"
  + It is a string of 5 characters with any character, which is the original, followed by any character, the original, any character, and the original such as `abaca`
* "(.)(.)(.).*\\3\\2\\1"
  + It is a group of three various characters followed by zero or more other characters, then the third character, second character, and first character such as `abcdefcba` or `abccba`
  
```{r}
teststring <- c("aaah", "noon", "mama", "rarer", "racecar")
# 3.a
str_view(teststring, "(.)\\1\\1")
# 3.b
str_view(teststring, "(.)(.)\\2\\1")
# 3.c
str_view(teststring, "(..)\\1")
# 3.d
str_view(teststring, "(.).\\1.\\1")
# 3.e
str_view(teststring, "(.)(.)(.).*\\3\\2\\1")

```
  

### Problem 4

Construct regular expressions to match words that:

* Start and end with the same character.
  + `"^(.).*\\1$"`
* Contain a repeated pair of letters (e.g. "church" contains "ch" repeated twice.)
  + `"(..).*\\1"`
  + Alternatively, `"([a-z][a-z]).*\\1"` to specify characters are letters
* Contain one letter repeated in at least three places (e.g. "eleven" contains three "e"s.)
  + `"(.).*\\1.*\\1"`
  + Alternatively, `"([a-z]).*\\1.*\\1"` to specify characters are letters
  
```{r}
teststring2 <- c("Mississippi", "noon", "mama", "rarer", "racecar")
# 4.a
str_view(teststring2, "^(.).*\\1$")
# 4.b
str_view(teststring2, "(..).*\\1")
# 4.c
str_view(teststring2, "(.).*\\1.*\\1")
```


  
  
